Four decades of water quality change in the upper San Francisco Estuary

### 2017-01-09 sftrends_manu.Rnw
map update with md10

The goal of this study was to provide a comprehensive description of nutrient trends in the Delta that will inform understanding of eutrophication dynamics and potential causes of water quality change in the larger Bay. We applied the newly-adapted method of weighted regression for tidal waters to describe nitrogen trends in  different spatial and temporal contexts. The specific objectives were to \begin{inparaenum}[1\upshape)]  
\item quantify and interpret trends over four decades at nine stations in the Delta, including annual, seasonal, and spatial changes in nitrogen analytes and response to flow variation,
\item provide detailed descriptions of three case studies in the context of conceptual relationships modelled with \ac{wrtds}.
\end{inparaenum}
The second objective evaluated three specific water quality stations in the Delta to demonstrate complexities with nutrient response to flow, effects of \ac{wwtp} upgrades on water quality, and effects of biological invasion by benthic filter feeders on primary production. Although quantitative descriptions of change can be ends in themselves, the results were expected to have greater impact as a means to more detailed understanding of ecosystem properties.  Products derived from \ac{wrtds} can be used to inform additional analyses, such as water quality response after removing annual, seasonal, or flow effects. Overall, this analysis is expected to further an ecological and management-based understanding of dynamics in San Francisco Bay, with implications for water quality restoration and protection of this prominent system.

>>>>

The goal of this study was to provide a comprehensive description of nutrient trends in the Delta that will inform understanding of eutrophication dynamics and potential causes of water quality change in the larger Bay. We applied the newly-adapted method of weighted regression for tidal waters to describe nitrogen trends in  different spatial and temporal contexts. The specific objectives were to \begin{inparaenum}[1\upshape)]  
\item quantify and interpret trends over four decades at ten stations in the Delta, including annual, seasonal, and spatial changes in nitrogen analytes and response to flow variation,
\item provide detailed descriptions of three case studies in the context of conceptual relationships modelled with \ac{wrtds}.
\end{inparaenum}
The second objective evaluated three specific water quality stations in the Delta to demonstrate complexities with nutrient response to flow, effects of \ac{wwtp} upgrades on water quality, and effects of biological invasion by benthic filter feeders on primary production. Although quantitative descriptions of change can be ends in themselves, the results were expected to have greater impact as a means to more detailed understanding of ecosystem properties.  Products derived from \ac{wrtds} can be used to inform additional analyses, such as water quality response after removing annual, seasonal, or flow effects. Overall, this analysis is expected to further an ecological and management-based understanding of dynamics in San Francisco Bay, with implications for water quality restoration and protection of this prominent system.
